### 2010-08-31 trunk/1tp/paper.tex
commit inicial con un template

%
% Latex example file for postgrads (04/10/03)
%
% If you have questions please email me: 
%
% M.L.Balogh@durham.ac.uk
%
% or find me in room OC312.
%
% you can copy this file and others from :
%
% http://star-www.dur.ac.uk/~balogh/latex/
%
% Note you can get Latex style files etc. from http://www.ctan.org

% 2. The aastex class file includes some useful astronomy definitions,
% for example special math symbols that are commonly used.  Use
% emulateapj instead to make Apj format.
% YOU SHOULD REMOVE THE TABLE OF
% CONTENTS PAGE WHEN USING APJ FORMAT.
%
%\documentclass[11pt,a4paper]{aastex}
\documentclass[11pt,a4paper]{emulateapj}
\bibliographystyle{apj}

%define general packages
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{natbib}
% spanish packages
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}

%internal short cuts
\def \HgA {H$\gamma_A$}
\def \gon {Gonz\'{a}lez}
\def \Hbp {H$\beta ^\prime$}

\begin{document}

\title{Diseño de reactores nucleares \HgA}
\author{Harald Kuntschner \& Michael Balogh}
\date{\today}

\begin{abstract}
This is a sample document which demonstrates some of the basic features
of \LaTeX.  You can easily reformat it for different document or bibliography styles.
\end{abstract}

\maketitle

%Section heading
\section{Introduction}
This file is an example of how to use \LaTeX.  In principle, you only have to worry
about the text content; formatting is taken care of by the options and arguments of
the ``$\backslash$documentclass'' command (other style files are available on the Web;
see \S\ref{sec-webrefs}).  In practice, however, things are not so simple.  Things like
figures and bibliographies can go all wrong when you change styles, so expect to have
to fiddle!

\section{Examples}

This is an example of a \LaTeX\ document.

\medskip

Galaxies are one of the \textbf{grandest} structures of the universe.
Their sizes range from as small as $10^6 \text{M}_\odot$ to as much as
$10^{13} \text{M}_\odot$, being built of stars, interstellar gas, dust
and perhaps dark matter. Yet galaxies were recognized only early in
this century (1920s) as \textit{vast} assemblages of stars, distinct
from our own Galaxy. Edwin~P.~Hubble (1889~-~1953) was one of the
pioneers in this field. He introduced a classification diagram for
galaxies, now widely known as the Hubble ``tuning fork'' diagram
\cite[][see Figure~\ref{fig:hubble_dia} for a revised Hubble
diagram]{hub36}.

%
% Subsection heading
%
\subsection{Example of a list and how to run \LaTeX}

% list environment
\begin{enumerate}

\item Write document: emacs test.tex $\Longrightarrow$ test.tex

\item \LaTeX\ document: latex test.tex $\Longrightarrow$ test.dvi

\item Create bibliography: bibtex test.tex $\Longrightarrow$ test.bbl

\item Correct X-refs: redo step 2. twice more!

\item View document: xdvi test.dvi

\item Produce ps file: dvips test.dvi $\Longrightarrow$ test.ps

\item Print document: lpr test.ps

\end{enumerate}

%
% Another subsection
%
\subsection{Math stuff}

For the following discussion let $G(n)$ be the spectrum of a galaxy
whose redshift and velocity dispersion are to be determined and $S(n)$
the spectrum of a {template} star at zero redshift, similar spectral type
to the galaxy and at instrumental resolution. The spectra have been
continuum subtracted and end-masked, and are each sampled in N bins of
equal $\log \lambda$ spacing. A velocity shift is then a linear
function of the spacing $\Delta \log \lambda$, with velocity increments
$\Delta v$ as follows:

%
% ``Pure'' equations
%
\begin{eqnarray}
  \Delta v &=& c \times \Delta \ln \lambda \\
           &=& c \times \ln (10) \times \Delta \log \lambda
\end{eqnarray}
where c is the speed of light. Now the normalized cross correlation
function is defined as:

\begin{equation}
  \label{equ:cross_corr}
  c(n)=G(m) \otimes S(n) = \frac{1}{N\sigma_G\sigma_S}\sum_{m=0}^{N}G(m)S(m-n)
\end{equation}
where $\sigma_G$ and $\sigma_S$ are the RMS deviations:

\begin{eqnarray}
  \sigma_G^2 &=& \frac{1}{N}\sum_{n=0}^{N}G(n)^2 \\
  \sigma_S^2 &=& \frac{1}{N}\sum_{n=0}^{N}S(n)^2 
\end{eqnarray}

%
%
\subsection{Table environment}
%
%

%
% Table environment
%
\begin{table}
  \caption{A Collection of Distance Measurements to Fornax}
  \label{tab:distance}
  \begin{center}
    \leavevmode
    \begin{tabular}{lll} \hline \hline              
  $m-M$          & Method              & Reference      \\ \hline 
  31.31$\pm$0.27 & Cepheid             & \citet{sil98}  \\
  31.22$\pm$0.06 & I-band SBF          & \citet{jen98}  \\
  31.32$\pm$0.24 & K$^\prime$-band SBF & \citet{jen98}  \\
  30.94$\pm$0.33 & I-band TFR          & \citet{bure96} \\
  31.07$\pm$0.13 & V-band GCLF         & \citet{Koh96}  \\
  30.85$\pm$0.11 & I-band GCLF         & \citet{Koh96}  \\
  31.14$\pm$0.14 & PNLF                & \citet{mcm93}  \\ \hline
  \multicolumn{3}{l}{}                                             \\       
  \multicolumn{3}{l}{SBF: Surface Brightness Fluctuations}         \\
  \multicolumn{3}{l}{TFR: Tully-Fisher relation}                   \\
  \multicolumn{3}{l}{PNLF: Planetary Nebulae Luminosity Functions} \\
  \multicolumn{3}{l}{GCLF: Globular Cluster Luminosity Functions}  \\
    \end{tabular}
  \end{center}
\end{table}

%
%
\subsection{Fonts and character sizes}
%
%
One can typeset a text {\sffamily\bfseries\large in a large sans serif
  bold typeface.} Or one can use a {\slshape\tiny tiny slanted
  typeface.} For more information see ``The \LaTeX\ Companion'' page
170 .

%
%
\subsection{\LaTeX\, info}\label{sec-webrefs}
%
%
\subsubsection{WEB info on \LaTeX}
There are links to some useful websites and \LaTeX\ resources are available
from: \\ {\tt http://star-www.dur.ac.uk/$\sim$balogh/latex/}.

\subsubsection{Books and guides for \LaTeX}
\begin{itemize}
\item {\it The LaTeX Companion}, by Michel Gossens, Frank Mittelbach, Alexander Samarin.
\item {\it LaTeX: A Document Preparation System -- User's Guide and
    Reference Manual}, by Leslie Lamport, 2nd ed
\item {\it The Not So Short Introduction to LaTeX2e}, Tobias Oetiker,
  see Starlink WEB page
\end{itemize}
\subsection{Bibilographies}

Bibliographies can be greatly facilitated with \LaTeX.  The best option is to keep a database
of references: see the file http://star-www.dur.ac.uk/$\sim$balogh/latex/example.bib for an example.
Use the apropriate bibliography style file (*.bst) to format the bibliography; read the first few
lines of this file for a description of usage.  Running ``bibtex'' on your document (see below)
will generate a *.bbl file with the appropriate references.  Alternatively, you can enter the
references directly, as described in the ``Not-so-short Guide''.

Here are some examples of the different ways you can reference a source with the
{\it natbib.sty} style file.  See the header of that file for more options.
\begin{itemize}
\item A first attempt to model this was made by \citet{bar96}.
\item This did not agree with earlier models \citep{bar96}.
\item There are several models for this phenomenon \citep{bar96,bau96}
\item There are several models for this phenomenon \citep[e.g., ][]{bar96,bau96,bure96}
\item Our model is better \citep[Fig. 5]{bau96}.
\item I want to do the references myself! (Baugh et al. 1996)\nocite{bau96}.
\end{itemize}
%
% References
%
\bibliography{example}

\end{document}